---
title: "Fundamentals of Computational Mathematics - Homework 1"
author: "Jered Ataky"
date: "2/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem

**Image transformation:**

Build the first letters for both your first and last name using point plots in R. 

Then, write R code that will left multiply (%>%) a square matrix (x) against each of the vectors of points (y).

Use a loop that changes the transformation matrix incrementally to demonstrate 1) shear, 2) scaling, 3)  rotation , and 4) projection in animated fashion.  

Hint:  Use x11() to open a new plotting window in R.



## Solution

**I am going to construct "A N" which is my last and middle name initial**


```{r}

# Building the letters A and N

x=c(seq(-1.5,-1.0, length.out = 500),seq(-1.25,-0.75,length.out=1000), seq(-0.5, -1, length.out = 500), 
    rep(0.5, 500), seq(0.5, 1.5, length.out = 1000), rep(1.5, 500))

y=c(seq(-1,1,length.out=500),rep(-0.2,1000), seq(-1,1,length.out=500), 
    seq(-1, 1, length.out = 500), seq(1, -1, length.out = 1000), seq(-1, 1, length.out = 500))

z=rbind(x,y)

plot(y~x, xlim=c(-3,3), ylim=c(-3,3))

```



```{r}

library(animation)

# Animating the initials 


myani = ani.record(reset = TRUE, replay.cur = FALSE)

a = diag(2)

x11()

for (i in seq(-3,3, length.out = 100)) {
  a[1,1] = i
  newmat = apply(z, 2, function(x) a%*%x)
  plot(newmat[2, ]~newmat[1, ], xlim =c(-3,3), ylim = c(-3,3))
  ani.record()
  
}


a = diag(2)

for (i in seq(-3,3, length.out = 100)) {
  a[2,1] = i
  newmat = apply(z, 2, function(x) a%*%x)
  plot(newmat[2, ]~newmat[1, ], xlim =c(-3,3), ylim = c(-3,3))
  ani.record()
  
}


a = diag(2)

for (i in seq(-3,3, length.out = 100)) {
  a[1,2] = i
  newmat = apply(z, 2, function(x) a%*%x)
  plot(newmat[2, ]~newmat[1, ], xlim =c(-3,3), ylim = c(-3,3))
  ani.record()
  
}


a = diag(2)

for (i in seq(-3,3, length.out = 100)) {
  a[2,2] = i
  newmat = apply(z, 2, function(x) a%*%x)
  plot(newmat[2, ]~newmat[1, ], xlim =c(-3,3), ylim = c(-3,3))
  ani.record()
  
}


myf = function(x)(matrix(c(cos(x), sin(x), -sin(x), cos(x)), byrow = TRUE, nrow = 2))

for (i in seq(0, 20, length.out = 100)) {
  
  a = myf(i)
  newmat = apply(z, 2, function(x) a%*%x)
  plot(newmat[2, ]~ newmat[1, ], xlim = c(-3,3), ylim = c(-3,3))
  ani.record()
  
  
}

```




