---
title: "Fundamentals of Computational Mathematics - Homework 10"
author: "Jered Ataky"
date: "4/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy).
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).
(c) Which strategy gives Smith the better chance of getting out of jail?

## Solution

### Part a

Here, the Markov chain would represent Smith's money evolution.

We can calculate it as:

$$P_{j} \ =\ \frac{1-\left(\frac{q}{p}\right)^{j}}{1-\left(\frac{q}{p}\right)^{m}}$$

```{r}
p <- 0.4
q <- 0.6
j <- 1
m <- 8

for (i in j:m){
  
  prob <- round(((1 - (q/p)^i) / (1 - (q/p)^m)), 4)

  print(paste0("For state ", i, ", the probability for timid srategy is ", prob)) 
  
}



```

We can see that the probability for smith with his 1 dollar is 0.0203

**Or as we know the state is 1 we can simply calculate it directly:**

```{r}
p <- 0.4
q <- 0.6

j <- 1
m <- 8

prob <- round(((1 - (q/p)^j) / (1 - (q/p)^m)), 4)

paste0("The probability for timid srategy is ", prob)


```



### Part b

I am going to use 3 different methods in this part (just for practice):

#### Trivial method

Smith has 1 dollar, he must then win 3 times in a row.

```{r}
p <- 0.4
n <- 3

prob <- p^n

paste0("The probability for bold strategy is ", prob)

```


#### Markov chain


We should here be concerned about the output of 1, 2, 4, 8 dollars:

```{r}

gain_prob <- matrix(c(1, .6, .6, .6, 0, 0, 0, 0, 0, 0, 0, .4, 0, 0, 0, 0, 0, .4, 0, 0, 0, 0, 0, .4, 1), nrow = 5)
gain_prob
```

Calculate the probability:

```{r}
init <- matrix(c(0, 1, 0, 0, 0), nrow = 1) # Initial state

prob1 <- init %*% gain_prob
prob2 <- prob1 %*% gain_prob
prob3 <- prob2 %*% gain_prob


paste0("The probability for bold strategy is ", prob3[5])

```


#### Binomial distribution

Smith needs 3 successes for the 3 trials to double the money until reaching 8 dollars

```{r}

prob <- dbinom(n, n, p)

paste0("The probability for bold strategy is ", prob)

```



### Part c

The bold strategy gives Smith a better chance of getting out of jail.






