---
title: "Fundamentals of Computational Mathematics - Homework 8"
author: "Jered Ataky"
date: "3/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 11 page 303

A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the first of these bulbs to burn
out? (See Exercise 10.)

### Solution

$$

 \begin{array}{l}
Let\ recall\ Exercise\ 10\ ( see\ page\ 303) :\\
\\
I\ am\ going\ to\ show\ that\ the\ density\ for\ M\ i\ exponential\ with\ mean\ \mu /n\ and\ then\ \\
use\ it\ to\ solve\ Exercise\ 11\\
\\
P( min( X_{1} ,...,\ X_{n}  >x) \ =\ ( P( X_{1}  >x))^{n} \ \\
=\left( e^{-\frac{x}{\mu }}\right)^{n} =e^{-\left(\frac{n}{\mu }\right) x}\\
Finding\ the\ min\ of\ this\ function:\ p_{min}( x_{1} ,...,x_{n}) \ =\ \frac{n}{\mu } e^{-\left(\frac{n}{\mu }\right) x}\\
\\
It\ is\ an\ exponential\ density\ with\ mean\ \frac{\mu }{n} \ \ ( 1)\\
\\
Using\ ( 1) \ to\ solve\ exercise\ 10,\ we\ have:\\
\\
The\ expected\ time\ for\ the\ first\ of\ these\ bulbs\ to\ burn\ out\ is\ E( minX_{i}) \ =\ \frac{1000}{100} \ =\ 10.\\
Thus\ the\ expected\ time\ is\ 10\ hours.\\
\end{array}

$$



## Problem 14 page 303

Assume that X1 and X2 are independent random variables, each having an
exponential density with parameter λ. Show that Z = X1 − X2 has density
fZ(z) = (1/2)λe−λ|z|




### Solution

$$

 \begin{array}{l}
X_{1} \ and\ X_{2} \ each\ having\ exponential\ density\ with\ parameter\ \lambda ,\ therefore:\\
\\
f_{X_{1}}( x) \ =\ f_{X_{2}}( x) =\ e^{\lambda x} \ if\ -\infty < x\leqslant 0;\ 0\ otherwise\ \ \\
\\
( see\ exercise\ 11for\ detail\ calculation\ on\ exponential\ density)\\
\\
\\
Z\ =\ X_{1} \ -\ X_{2} \ =\ X_{1} \ +\ ( -X_{2})\\
\ \ \ \  =\ X_{1} \ +\ X^{'}_{2} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \left( X^{'}_{2} \ =\ -X_{2}\right)\\
\\
Z\ =\ X_{1} +\ X^{'}_{2} \ \ will\ have\ a\ density:\\
\\
f_{z}( z) \ =\int ^{\infty }_{0} \lambda e^{-\lambda x} \lambda e^{-\lambda ( x_{1} -z)} dx_{1} \ \ \ \ \ \ \ \ \ ( Convolution\ formula)\\\\ \ \ \ \ \ \ \ \ \ \ 
=\ \int ^{\infty }_{-\infty } \lambda ^{2} e^{-2\lambda x_{1}} e^{\lambda z} dx_{1} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\\ \
\\
\ \ \ \ \ \ \ \ \ \ =\int ^{\infty }_{-\infty } \lambda ^{2} e^{\lambda ( z-2x_{1})} dx_{1} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
\\
This\ can\ be\ generalized\ as:\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
f_{z}( z) \ =\ \int ^{\infty }_{0} e^{\lambda ( z-2x_{1})} dx_{1} \ =\ \frac{1}{2\lambda } e^{\lambda z} ,\ z< 0;\ \ \ \ \ \ \ \ \ \\\\\
\ \ \ \ \ \ \ \ \  =\ \int ^{\infty }_{0} e^{\lambda ( z-2x_{1})} dx_{1} \ =\ \frac{1}{2\lambda } e^{\lambda z}\left( e^{-2\lambda z}\right) =\frac{1}{2\lambda } e^{-\lambda x} ,\ z\geqslant 0\ \ \ \ ( 2) \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
\\
Thus,\ from\ ( 1) \ and\ ( 2) ,\ we\ have:\\
\\
f_{z}( z) \ =\frac{1}{2\lambda } e^{-\lambda |z|} \ \ \ \ \ \ \ \ \ \\
\end{array}

$$


## Problem 1 page 330

Let X be a continuous random variable with mean µ = 10 and variance
σ2 = 100/3. Using Chebyshev’s Inequality, find an upper bound for the
following probabilities.

(a) P(|X − 10| ≥ 2).
(b) P(|X − 10| ≥ 5).
(c) P(|X − 10| ≥ 9).
(d) P(|X − 10| ≥ 20).


### Solution

$$\ Let\ define\ Chebyshev’s\ Inequality:\ P( |X\ −\mu |\geqslant k\sigma ) \ \leqslant \frac{1}{k^{2}}\\$$

```{r}
sd <- sqrt(100/3)

```


(a) P(|X − 10| ≥ 2)

```{r}

mu <- 10
k <- 2 / sd

upper_bound <- round(1 / (k**2), 4)

paste0("The upper bound is : ", upper_bound)
```

(b) P(|X − 10| ≥ 5).

```{r}

mu <- 10
k <- 5 / sd

upper_bound <- round(1 / (k**2), 4)

paste0("The upper bound is : ", upper_bound)
```

(c) P(|X − 10| ≥ 9).

```{r}

mu <- 10
k <- 9 / sd

upper_bound <- round(1 / (k**2), 4)

paste0("The upper bound is : ", upper_bound)
```


(d) P(|X − 10| ≥ 20).

```{r}

mu <- 10
k <- 20 / sd

upper_bound <- round(1 / (k**2), 4)

paste0("The upper bound is : ", upper_bound)
```

```{r}
library(matlib)



```

