---
title: "Fundamentals of Computational Mathematics - Homework 3"
author: "Jered Ataky"
date: "2/13/2021"
output:
  pdf_document: default
  html_document:
    number_sections: yes
  openintro::lab_report: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PROBLEM SET 1

### (1) The rank of matrix A

$$
\begin{gather*}
I\ am\ going\ to\ use\ RREF\ to\ find\ the\ rank\ of\ A\\
\\
A\ =\begin{bmatrix}
1 & 2 & 3 & 4\\
-1 & 0 & 1 & 3\\
0 & 1 & -2 & 1\\
5 & 4 & -2 & -3
\end{bmatrix} R_{2}\rightarrow R_{3} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
=\begin{bmatrix}
1 & 2 & 3 & 4\\
0 & 1 & -2 & 1\\
-1 & 0 & 1 & 3\\
5 & 4 & -2 & -3
\end{bmatrix} R_{3} =R_{1} \ +\ R_{3} ;\ R_{4} \ =-5R_{1} \ +\ R_{4} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
\ \ \ =\begin{bmatrix}
1 & 2 & 3 & 4\\
0 & 1 & -2 & 1\\
0 & 2 & 4 & 7\\
0 & 6 & -17 & -23
\end{bmatrix} R_{3} =-2R_{2} +R_{3} ;\ R_{4} =-6R_{2} +R_{4} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
\\
=\begin{bmatrix}
1 & 2 & 3 & 4\\
0 & 1 & -2 & 1\\
0 & 0 & 8 & 5\\
0 & 0 & -5 & -29
\end{bmatrix} R_{3} =\frac{1}{8} R_{3} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
=\begin{bmatrix}
1 & 2 & 3 & 4\\
0 & 1 & -2 & 1\\
0 & 0 & 1 & \frac{5}{8}\\
0 & 0 & -5 & -29
\end{bmatrix} R_{4} =5R_{3} +R_{4} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
=\begin{bmatrix}
1 & 2 & 3 & 4\\
0 & 1 & -2 & 1\\
0 & 0 & 1 & \frac{5}{8}\\
0 & 0 & 0 & -\frac{207}{8}
\end{bmatrix} R_{4} =-\frac{8}{207} R_{4} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ =\begin{bmatrix}
1 & 2 & 3 & 4\\
0 & 1 & -2 & 1\\
0 & 0 & 1 & \frac{5}{8}\\
0 & 0 & 0 & 1
\end{bmatrix} \ R_{3} =-\frac{5}{8} R_{4} \ +R_{3} ;\ R_{2} =-R_{4} +R_{2} ;\ R_{1} =-4R_{4} \ +R_{1} \ \ \ \ \ \ \ \ \ \ \\
\\
\\
=\begin{bmatrix}
1 & 2 & 3 & 0\\
0 & 1 & -2 & 0\\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1
\end{bmatrix} \ R_{2} =2R_{3} +R_{2} ;\ R_{1} =-3R_{3} +R_{1} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
=\begin{bmatrix}
1 & 2 & 0 & 0\\
0 & 1 & 0 & 0\\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1
\end{bmatrix} \ R_{1} =- 2R_{1} +R_{1} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
=\begin{bmatrix}
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0\\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1
\end{bmatrix} \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
As\ we\ can\ see\ the\ RREF\ of\ A\ with\ leading\ 1s,\ rank( A) \ =\ 4\\
\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
\end{gather*}

$$


Let verify the above solution using rankmatrix

```{r}
library(Matrix)

A <- matrix(c(1,-1,0,5, 2,0,1,4, 3,1,-2,-2, 4,3,1,-3), nrow = 4)
R <- rankMatrix(A)

message("The rank of A is ", R)


```



### (2) Maximum and minimum of an mXn matrix

Given an mXn matrix where m > n, the maximum rank can be **n** 
since the rank has to be no greater than the smaller of the row or column
dimension.

Assuming that the matrix is non-zero, the minimum rank can be **1**
since the rank of a matrix would be zero only if the matrix had no elements.


### (3) Rank of matrix B

$$
\begin{gather*}
B=\begin{bmatrix}
1 & 2 & 1\\
3 & 6 & 3\\
2 & 4 & 2
\end{bmatrix}\\
\\
Looking\ at\ matrix\ B,\ R_{2} \ =\ 3R_{1} \ and\ R_{3} \ =\ 2R_{1} ,\\
all\ rows\ vectors\ are\ multiple\ of\ the\ same\ nonzero\ vector\ and\ det( B) \ =\ 0.\\
That's\ it,\ there\ is\ one\ and\ only\ one\ linearly\ independent\ row.\\
\\
Therefore,\ rank( B) \ =\ 1.\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
\end{gather*}

$$



Let verify the above solution using rankmatrix

```{r}

B <- matrix(c(1,3,2, 2,6,4, 1,3,2), nrow = 3)
R <- rankMatrix(B)

message("The rank of B is ", R)


```


## PROBLEM SET 2

**Eigenvalues and Eigenvectors**


$$
\begin{gather*}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \\
\\
\\
\\
\\
A\ =\begin{bmatrix}
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6
\end{bmatrix}\\
\\
1.\ Eigenvalues\\
\\
The\ polynomial\ characteristic\ is\ given\ by:\\
\\
P( \lambda ) \ =det( A-\lambda I) \ =\ det\left(\begin{bmatrix}
1-\lambda  & 2 & 3\\
0 & 4-\lambda  & 5\\
0 & 0 & 6-\lambda 
\end{bmatrix}\right)\\
\\
using\ column\ 1\ to\ calculate\ the\ determinant:\\
\\
det\left(\begin{bmatrix}
1-\lambda  & 2 & 3\\
0 & 4-\lambda  & 5\\
0 & 0 & 6-\lambda 
\end{bmatrix}\right) \ =( 1-\lambda )\begin{vmatrix}
4-\lambda  & 5\\
0 & 6-\lambda 
\end{vmatrix}\\
\\
=\ ( 1-\lambda )( 4-\lambda )( 6-\lambda )\\
\\
P( \lambda ) \ =\ 0\ \Longrightarrow \ \lambda =1,\ \lambda =4,\ \lambda =6\\
\\
The\ eigenvalues\ are\ \lambda =1,\ 4,\ 6\ with\ algebraic\ multiplicities\ \\
\alpha ( 1) \ =\ 1,\ \alpha ( 4) \ =\ 1,\ and\ \alpha ( 6) \ =\ 1.\\
\\
2.\ Eigenvectors\\
\\
For\ \lambda =1:\\
\\
A-I\ =\ \begin{bmatrix}
0 & 2 & 3\\
0 & 3 & 5\\
0 & 0 & 5
\end{bmatrix}\rightarrow \ v_{3} =v_{2} =0\ and\ v_{1} \ =\ v_{1}\\
\\
\Longrightarrow \epsilon ( 1) \ =\left\{\begin{bmatrix}
1\\
0\\
0
\end{bmatrix}\right)\\
\\
For\ \lambda =4:\\
\\
A-4I\ =\ \begin{bmatrix}
-3 & 2 & 3\\
0 & 0 & 5\\
0 & 0 & 2
\end{bmatrix}\rightarrow \ v_{3} =0\ and\ -3v_{1} \ +\ 2v_{2} \ =0\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \rightarrow \ v_{1} \ =\ \frac{2}{3} v_{2} \ \ \ \ \ \ \ \\
\\
\ \ \ \ \Longrightarrow \epsilon ( 4) \ =\left\{\begin{bmatrix}
2\\
3\\
0
\end{bmatrix}\right\}\\
\\
For\ \lambda =6:\\
\\
A-6I\ =\begin{bmatrix}
-5 & 2 & 3\\
0 & -2 & 5\\
0 & 0 & 0
\end{bmatrix}\rightarrow v_{2} =\frac{5}{2} v_{3} \ and\ 5v_{1} \ =\ 8v_{3}\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \rightarrow \ v_{1} =\frac{8}{5} v_{3} \ and\ v_{2} =\frac{5}{2} v_{3}\\
\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \Longrightarrow \ \epsilon ( 6) \ =\left\{\begin{bmatrix}
16\\
25\\
10
\end{bmatrix}\right\}\\
\\
The\ eigenvectors\ yield\ the\ geometric\ multiplicities\ \gamma ( 1) \ =1,\ \gamma ( 4) \ =1,\ \gamma ( 6) \ =1\\
\\
\ \ \ \ \ \\
\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
\end{gather*}


$$