---
title: "Fundamentals of Computational Mathematics - Homework 5"
author: "Jered Ataky"
date: "2/28/2021"
output:
  pdf_document: default
  html_document:
    number_sections: yes
  openintro::lab_report: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: Bayesian


```{r}

sens <- 0.96
spec <- 0.98

p_rate <- 0.001

p_desease <- round(((sens * p_rate) / ((sens * p_rate) + (1 - spec)*(1 - p_rate))), 4)


paste0("The probability that an individual who is reported as positive by the new test actually has disease is ", p_desease)

median_cost <- 100000
test_cost <- 1000
total_ind_treated <- 100000

# First year cost for treating 100,000 individuals:
# Considering that each positive individual (individual to be treated) did the test
# And each individual needs to pay $100,000 for the treatment 

total_test_cost <- (median_cost + test_cost) * total_ind_treated # But some patients would be false positive (sens = 96%                                                                     not 100%)

paste0("The total cost will be $", total_test_cost)

```


## Problem 2: Binomial

```{r}

# X = number of inspection
p <- 0.05
n <- 24

a <- round(dbinom(2, 24, 0.05), 4)

paste0("P(X = 2) = ", a)


b <- round(1 - pbinom(1, 24, 0.05), 4)

paste0("Probabibility after 24 months you receive 2 or more inspections is ", b)



c <- round(pbinom(1, 24, 0.05), 4)

paste0("Probabibility after 24 months you receive fewer than 2 is ", c)


# Expected value

u <- n*p
paste0("The expected value is ", u)

# Standard deviation

s <- round(sqrt(n*p*(1-p)), 4)
paste0("The standard deviation is ", s)

```


## Problem 3: Poisson distribution

```{r}

lambda<- 10
t <- 1

# Exactly 3 per hour

a <- dpois(3, 10)
paste0("P(X = 2) = ", a)

b <- 1 - ppois(10, 10)
b

# Expected value and std

t <- 8

u <- lambda*t
paste0("We will expect ", u, " patients to arrive in 8 hours")

s <- round(sqrt(lambda*t), 2)
paste0("The standard deviation for patients to arrive in 8 hours is ", s)

```

Let now consider each family provider clinic is open 8hrs a day:

From the above calculation, the clinic expects 80 patients a day.

With 3 family practice providers that can see 24 templated patients each day,
the total number of patients seen would be 72 which is less than the expected
number of patients per day. The demand is way more than the supply,
thus, the percent utilization > 100%.

As the clinic can only see 90% of patients that arrive,
I would recommend the clinic to hire another family providers. With this, they
will serve all the patients and reduce the percent utilization as well.

If the clinic can't afford t hire another family provider,
the clinic could to start introducing an "appointment only" visit. 
But With this, the clinic won't be able to serve
all the patients and the receptionists will get busier scheduling appointment.
The clinic might need to hire more receptionists to handle phone calls
which cost in the end can be as same as hiring another family provider.



## Problem 4: Hypergeometric 

```{r}

nurses <- 15
non_nurses <- 15

# Let nt be number of nurses sent to trip, n_nt number of non nurses sent to trip

n <- 6 # total for the trip
nt <- 5
n_nt <- 1

p <- round(dhyper(5, 15, 15, 6), 4)

paste0("The probability the subordinate would have selected fives is ", p)

# Let en be expected value for nurse, n_en expected value for non-nurses

en <- n*(nurses/(nurses + non_nurses))

n_en <- n*(non_nurses/ (nurses + non_nurses))

paste0(" We would have expected the subordinate to send ", en, " nurses and ", n_en, " non-nurses" )



```



## Problem 5: Geometric distribution 

```{r}

p <- 0.001
n <- 1200

# For probability that the driver will be seriously injured during the course of the year,
# I would consider it as a geometric distribution where x <= 1200 hours
# in other word: During the course of year: continuously from 0 to 1200 hours
# as compared to at 1200 hours which I would have used dgeom to find the probability

a <- round(pgeom(n, p), 4)

paste0("The probability that the driver will be seriously injured during the course of the year is ", a)


# Same analogy but 15 months = 1500 hours
n1 <- 1500

a1 <- round(pgeom(n1, p), 4)

paste0("The probability that the driver will be seriously injured during the course of 15 months is ", a1)

```
**Plotting my analysis**

```{r}

library(dplyr)
library(ggplot2)

data.frame(x = seq(0, 1500, 100), 
           pmf = dgeom(x = seq(0, 1500, 100), prob = p),
           cdf = pgeom(q = seq(0, 1500, 100), prob = p, lower.tail = TRUE)) %>%
  mutate(Failures = ifelse(x <= n, n, "pass a year")) %>%
  
ggplot(aes(x = factor(x), y = cdf, fill = Failures)) +
  geom_col() +
  
  geom_text(
    aes(label = round(cdf,2), y = cdf + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  
  labs( x = "Failures prior success", y = "probability", 
        title = "Cumulative Probability to 1500 hours for faillure.") 





```


**Expected value**

```{r}
p <- 0.001
n <- 1200

u <- 1/p

paste0("The expected number of hours that a driver will drive before being seriously injured is ", u)

```

**Conditional probability**

Geometric distribution is one of the two memoryless distributions.
Therefore, we have:

```{r}
# Using memoryless property, the 1200 hours don't affect the next 100 hours
# In the next 100 hours, P(X<x) = P(X,100)

n <- 100

a <- round(pgeom(n, p), 4)

paste0("The probability i ", a)


```


## Problem 6

```{r}

# Solve this problem as a binomial distribution
# P(X>x)

p = 0.001
n <- 1000

x <- 2

prob <-round(1 - pbinom(x, n, p), 4) 

paste0("the probability is ", prob)

u <- n*p

paste0("The expected value is ", u)


# Using poisson distribution

lambda <- 1

p_g <- round(1- ppois(x, lambda), 4)
p_g
u_pg <- lambda
u_pg

```

Both methods give us the same answer (very closer)


## Problem 7

```{r}
lower <- 0
upper <- 30

fx <- 1 / (upper - lower) # base

#a. Patient to wait more than 10 minutes, P(X>10)

x <- 10
prob10 <- round (fx * (upper - x), 4)

paste0("Probability for a patient to wait more than 10 minutes is ", prob10)


# The patient has already waited 10 minutes... P(X>5|X=10)

 
x1 <- 10 # new starting point

x <- x1 + 5 # Start waiting time

fx1 <- 1 / (upper - x1)  # new base

prob5 <- round((fx1 * (upper - x)), 4) # We can aslso use conditional probability formula

paste0("Probability  is ", prob5)

# Expected value

exp_v <- (lower + upper) / 2

paste0("The expected waiting time is ", exp_v)



```



## Problem 8

```{r}
u_exp <- 10 # expected value

rate <- 1/ u_exp # rate

# a. Expected failure time: exp_fail

exp_fail <- 1 / rate  

paste0("Expected failure time is ", exp_fail)


# Standard deviation

exp_std <- exp_fail

paste0("The std is ", exp_fail)

# More than 8 years P(x>8)

x <- 8

prob <- round(exp(-rate*x), 4) 

paste0("The probability that my MRI will fail after 8 years is ", prob)

# Exponential distribution has a memoryless property
# The first 8 years don't affect the MRI as it didn't break
# In the next two years = within two years = P(X<2)

y <- 2

prob2 <- round(1 - exp(-rate*y), 4) 

paste0("The probability is ", prob2)



```
