---
title: "Fundamentals of Computational Mathematics - Homework 8"
author: "Jered Ataky"
date: "3/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}

library(tidyverse)
```



## The data

I am going to do the summary of the data...

```{r}

summary(cars)

```


## Visualize the data

I'm going to visualize the data and see at first glance how it looks...



```{r}

plot(cars$speed, cars$dist, xlab = 'Speed (mph)', ylab = 'Stopping Distance (ft)',
     main = 'Stopping distance Vs Speed')


```

We can see that the data appear to follow some linearity...


Now let build the linear model


## Model


```{r}

lin_model <- lm(cars$dist~cars$speed)
summary(lin_model)

```


The correlation...

```{r}

cars %>%
  summarise(cor(dist, speed, use = "complete.obs"))

```



We can write the linear regression as:

dist = -17.5791 + 3.9324(speed)



Let now visualize the model...

```{r}

plot(cars$speed, cars$dist, xlab = 'Speed (mph)', ylab = 'Stopping Distance (ft)',
     main = 'Stopping distance Vs Speed')

abline(lin_model, col='red')


```



## Quality evaluation

There is also a fairly strong positive relationship between the stopping distance
and the speed as the correlation coefficient is 0.8069 but the r squared is 0.6511
meaning the model explains only 65.11 percent of data's variation.


## Diagnostics: Residual analysis



Let plot the residuals Vs Fitted Values...

```{r}

plot(lin_model$fitted.values, lin_model$residuals, xlab = 'Fitted Values', ylab = 'Residuals')

abline(0,0, col='red')


```


There is not an apparent pattern in the residuals plots. This indicates there is a linear
relationship between the two variables.

Let check now the normality of the residuals wih he histogram and the QQ plot.


Histogram...

```{r}

ggplot(data = cars, aes(x = lin_model$residuals)) + 
  geom_histogram() +
  xlab("Residuals")


```



Normal Q-Q plot


```{r}

qqnorm(lin_model$residuals)
qqline(lin_model$residuals)

```



With the correlation coefficient of 0.8069, r square of 0.6511, and looking at 
the above graphs, we can say that the linear model does a good job at explaining but
the speed as the only variable is not enough. I'd say that considering other factors or 
paremeters would make the model more robust.

































































