---
title: "Fundamentals of Computational Mathematics - Homework 12"
author: "Jered Ataky"
date: "4/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)


```


## Introduction

1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the
variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss
whether the assumptions of simple linear regression met.

2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06
power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r
re-run the simple regression model using the transformed variables. Provide and interpret the F
statistics, R^2, standard error, and p-values. Which model is "better?"

3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life
expectancy when TotExp^.06=2.5.

4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error,
and p-values. How good is the model?
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why
or why not? 

## Data Exploration

```{r}

raw_data <- read.csv('https://raw.githubusercontent.com/jnataky/Computational_Mathematics/main/Assignments/who.csv')

glimpse(raw_data)

```

```{r}
df <- raw_data

```


## 1. LifeExp~TotExp: Simple linear regression


```{r}

# Scatter plot
df %>%
  ggplot(aes(x = TotExp, y = LifeExp)) +
  geom_point()


```


```{r}

# Linear model

df_lm <- lm(data = df, LifeExp~TotExp)
summary(df_lm)


```

R square = 0.2577; The model explains 25.77% of variability of the response data (LifeExp) around its mean.
rse = 9.371; spread is lower
F-stat = 65.26 on 1 and 188 DF
p-value = 7.714e-14; Reject the null hypothesis


**Assumptions**


Le take a look on each variable and its residuals...


1. Normality 

```{r}

qqnorm(df_lm$residuals)
qqline(df_lm$residuals)


```


Residuals are not nearly normal.


2. Linearity

The scatter plot will have to check the linearity...

```{r}

# Scatter plot
df %>%
  ggplot(aes(x = TotExp, y = LifeExp)) +
  geom_point()


```

Looking at the above plot, we can say at first glace that the data don't look linear.



3. Constant variability

```{r}

ggplot(data = df_lm, aes(x = .fitted, y = .resid)) +
  geom_point()+
  
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")


```


The variability on the residuals is not constant.



**At the end, we can say that the assumptions of simple linear regression are not met**


## 2. Model transformed 

```{r}

# Transform the data

df$LifeExp <- df$LifeExp**4.6
df$TotExp <- df$TotExp**0.06

```



```{r}
# Re-run the model

df_lm2 <- lm(data = df, LifeExp~TotExp)
summary(df_lm2)

```



R square = 0.7298; The model explains 72.98% of variability of the response data (LifeExp) around its mean.
rse = 90490000; Larger deviation from the true regression line
F-stat = 507.7 on 1 and 188 DF
p-value = 2.2e-16; Reject the null hypothesis


Visualize the new model...

```{r}

# Scatter plot
df %>%
  ggplot(aes(x = TotExp, y = LifeExp)) +
  geom_point()


```

Residuals plot...

```{r}

qqnorm(df_lm2$residuals)
qqline(df_lm2$residuals)


```


The residuals plot shows there is more constant variability and also comparing different parameters(r square, F-Stat), I would say that the transformed model is better then the previous one.


## 3. Forecasting 

From 2:

LifeExp = -736527910 + (620060216)*(TotExp)

```{r}

TotExp2 <- 1.5
LifeExp11 <- -736527910 + (620060216)*(TotExp2)

paste0("Life expectancy when  TotExp^.06 = 1.5 is ", LifeExp11)

TotExp4 <- 2.5
LifeExp22 = -736527910 + (620060216)*(TotExp4)

paste0("Life expectancy when  TotExp^.06 = 2.5 is ", LifeExp22)


```

Transforming back LifeExp = LifeExp^(1/4.6)...

```{r}

LifeExp21 <- round(LifeExp11**(1/4.6), 0)

paste0("Life expectancy when  TotExp^.06 = 1.5 and  transorming back LifeExp is ", LifeExp21)


LifeExp41 <- round(LifeExp22**(1/4.6), 0)

paste0("Life expectancy when  TotExp^.06 = 1.5 and  transorming back LifeExp is ", LifeExp41)


```


## 4. Multiple regression model

Transform back the data...

```{r}

# Transform the data

df$LifeExp <- df$LifeExp**(1/4.6)
df$TotExp <- df$TotExp**(1/0.06)

```



```{r}

# Run the model

df_lm3 <- lm(data = df, LifeExp~PropMD + TotExp + PropMD*TotExp)
summary(df_lm3)

```


R square = 0.3574; The model explains 35.74% of variability of the response data (LifeExp) around its mean.
rse = 8.765; Less deviation from the true regression line
F-stat = 34.49 on 3 and 186 DF
p-value = 2.2e-16; Reject the null hypothesis


*Residuals analysis**


```{r}

qqnorm(df_lm3$residuals)
qqline(df_lm3$residuals)


```

```{r}

ggplot(data = df_lm3, aes(x = .fitted, y = .resid)) +
  geom_point()+
  
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")


```



Looking at the plots, we can see that there is not constant variability,
residuals are not really normal. Also we saw that r squared was small...
With all of these, we can say that the model doesn't fit well the data.


## 5. New forescating

LifeExp = 6.277e+01 + (1.497e+03)(PropMD) + (7.233e-05)(TotExp) + (-6.026e-03)(PropMD)*(TotExp)

```{r}

PropMD <- 0.03
TotExp <- 14

LifeExp <- round(6.277e+01 + (1.497e+03)*(PropMD) + (7.233e-05)*(TotExp) + (-6.026e-03)*(PropMD)*(TotExp), 0)

paste0("Life expectancy when PropMD = 0.3 and TotExp = 14 is ", LifeExp)

```

```{r}

max(df$LifeExp)


```

The forecast does not seem realistic because per WHO the max LifeEXp is 83 years.
108 years is way above the max



