---
title: "Fundamental of Computational Mathematics - Homework 6"
author: "Jered Ataky"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries

```{r}

library(DescTools)


```


## Question 1

A bag contains 5 green and 7 red jellybeans. How many ways can 5 jellybeans be
withdrawn from the bag so that the number of green ones withdrawn will be less than
2?

**Answer**

```{r}
g <- 5
r<- 7

## Order does not matter: combination

# Number of green be less than 2

# Case 1: number of green ones is 0 and red ones 5


total1 <- CombN(r, 5) # Just need to combine 7 reds ones 5 by 5


# Case 2: number of greens is 1 and reds 4

total2 <- CombN(g, 1) * CombN(r, 4) # Combine 5 greens ones by 1, and 7 reds one by 4

# Total case

total <- total1 + total2

paste0(" There is a total of ", total, " ways")

```

## Question 2

A certain congressional committee consists of 14 senators and 13 representatives. How
many ways can a subcommittee of 5 be formed if at least 4 of the members must be
representatives?


**Answer**


```{r}

sen <- 14
reps <- 13

## Order does not matter: combination

# At least 4 of members be reps: 2 cases

# Case 1: subcommittee of 4 reps and 1 senator

subcom1 <- CombN(reps, 4)*CombN(sen, 1)

# Case 2: subcommittee of 5 reps and 0 senator

subcom2 <- CombN(reps, 5)*CombN(sen, 0)

# Total

subcom <- subcom1 + subcom2

paste0("There is ", subcom, " ways of forming a subcommittee")


```


## Question 3

If a coin is tossed 5 times, and then a standard six-sided die is rolled 2 times, and finally
a group of three cards are drawn from a standard deck of 52 cards without
replacement, how many different outcomes are possible?


**Answer**

```{r}

# coin tossed 5 times

t1 <- CombN(2, 1)**5 # same as multiplying itself 5 times

# standard six-sided die is rolled 2 times

t2 <- CombN(6, 1)*CombN(6, 1)

# 3 cards drawn

t3 <- CombN(52, 3)

# Total

t <- t1*t2*t3


paste0("There are ", t, " different outcomes possible")


```


## Question 4

3 cards are drawn from a standard deck without replacement. What is the probability
that at least one of the cards drawn is a 3? Express your answer as a fraction or a
decimal number rounded to four decimal places.

**Answer**

```{r}
# 3 cards drawn.... order does not matter: combinaison

possible_outcome <- CombN(52, 3)

## At least one of 3 cards is 3
## 3 cases: one 3, two 3 or three 3
# Note that there are four 3 in a standard deck

n_outcome <- CombN(4,1)*CombN(48, 2) + CombN(4, 2)*CombN(48, 1) + CombN(4, 3)*CombN(48, 0)

prob <- round(n_outcome / possible_outcome, 4)


paste0("The probability that at least one of the cards drawn is a 3 is ", prob)

```


## Question 5

Lorenzo is picking out some movies to rent, and he is primarily interested in
documentaries and mysteries. He has narrowed down his selections to 17
documentaries and 14 mysteries.

Step 1. How many different combinations of 5 movies can he rent?

**Answer**

```{r}
docs <- 17
myst <- 14

# No condition on selection. Need just to add both type of movies then combine them by 5

sel <- CombN(docs + myst, 5)

paste0("Lorenzo can rent ", sel, " different combinations of 5 movies")

```

Step 2. How many different combinations of 5 movies can he rent if he wants at least
one mystery?

**Answer**

```{r}

# At least one mystery means 1, 2, 3, 4 or 5 mystery movie

sel <- CombN(myst, 1)*CombN(docs, 4) + CombN(myst, 2)*CombN(docs, 3) + CombN(myst, 3)*CombN(docs, 2) +
  CombN(myst, 4)*CombN(docs, 1) + CombN(myst, 5)*CombN(docs, 0)

paste0("Lorenzo can rent ", sel, " different combinations of 5 movies can he rent if he wants at least
one mystery")

```



## Question 6


In choosing what music to play at a charity fund raising event, Cory needs to have an
equal number of symphonies from Brahms, Haydn, and Mendelssohn. If he is setting up
a schedule of the 9 symphonies to be played, and he has 4 Brahms, 104 Haydn, and 17
Mendelssohn symphonies from which to choose, how many different schedules are
possible? Express your answer in scientific notation rounding to the hundredths place.

**Answer**

```{r}
 
brahms <- 4
haydn <- 104
mendel <- 17

## Cory needs to have an equal number of symphonies...
# This means 3 from each repertoire
# Order does not matter: combination of each repertoire by 3

total <- CombN(4, 3)*CombN(104, 3)*CombN(17, 3)

# Music can play at any order: permuting them

total <- total*factorial(9)

total <- format(total, scientific = TRUE, digits = 3)

paste0("There are ", total, " different schedules possible")

```


## Question 7

An English teacher needs to pick 13 books to put on his reading list for the next school
year, and he needs to plan the order in which they should be read. He has narrowed
down his choices to 6 novels, 6 plays, 7 poetry books, and 5 nonfiction books.

Step 1. If he wants to include no more than 4 nonfiction books, how many different
reading schedules are possible? Express your answer in scientific notation rounding to
the hundredths place.


**Answer**

```{r}
# Order matters: variation/permutation

perms <- function(n, r){
  return(factorial(n)/factorial(n - r))
}

novels <- 6
plays <- 6
poetry <- 7
n_f <- 5 # nonfiction books

n_t <- novels + plays + poetry # Adding other books

# No more than 4 nonfiction books means at most 4
# 5 possibilities: 0, 1, 2, 3, 4

# Different options for combining those books are given by:

poss <- CombN(n_f, 0)*CombN(n_t, 13) + CombN(n_f, 1)*CombN(n_t, 12) + CombN(n_f, 2)*CombN(n_t, 11)+
  CombN(n_f,3)*CombN(n_t, 10) + CombN(n_f, 4)*CombN(n_t, 9)



# The different combinations should be at any order: permuting them

poss_t <- poss*perms(13,13)

poss_t <- format(poss_t, scientific = TRUE, digits = 3)

poss <- format(poss, scientific = TRUE, digits = 3) # in scientific notation


paste0("There are ", poss_t, " different reading schedules possible")

```



Step 2. If he wants to include all 6 plays, how many different reading schedules are
possible? Express your answer in scientific notation rounding to the hundredths place.


**Answer**

```{r}

# Arrange other books by 7

n_s <- novels + poetry + n_f
t1 <- perms(n_s, 7)

# permutation of 13 books

t2 <- perms(13, 13) 

# Total ways

total <- t1*t2 # take the 13 election and arrange them in different order

total <- format(total, scientific = TRUE, digits = 3)

paste0("There are ", total, " different reading schedules possible")



```


## Question 8

Zane is planting trees along his driveway, and he has 5 sycamores and 5 cypress trees to
plant in one row. What is the probability that he randomly plants the trees so that all 5
sycamores are next to each other and all 5 cypress trees are next to each other?
Express your answer as a fraction or a decimal number rounded to four decimal places.


**Answer**

```{r}

syc <- 5
cyp <- 5

choose(2,1)*choose(5,5)*choose(5,0)/choose(10,5)

# Possible outcome: planting trees at any order

n<- 5 # number of trees

poss <- CombN(syc+cyp, n) 

# We need to choose one in a group of 2 types of trees
# Choose 5 in a group of 5
# Multiply by the number of trees

n_poss <- CombN(2, 1)*CombN(5, 5)*CombN(5, 0)

total_poss <- round(n_poss / poss, 4)

paste0("The probability is ", total_poss)


```


## Question 9

If you draw a queen or lower from a standard deck of cards, I will pay you $4. If not, you
pay me $16. (Aces are considered the highest card in the deck.)

Step 1. Find the expected value of the proposition. Round your answer to two decimal
places. Losses must be expressed as negative values.


**Answer**

```{r}

# card less or equal to queen: 11
num <- 11

num_t <- num*4 # total of cards lower or equal to queen

num1 <- 2 # num of cards greater then queen
num1_t <-num *4

total <- 52 #deck

exp_value <- round(((num_t / total)*4) - ((num1_t /total)*16), 2)

paste0("the expected value is $ ", exp_value)



```
If you draw a queen or lower from a standard deck of cards, I will pay you $4. If not, you
pay me $16. (Aces are considered the highest card in the deck.)


Step 2. If you played this game 833 times how much would you expect to win or lose?
Round your answer to two decimal places. Losses must be expressed as negative
values.

**Answer**

For 1 game, we are expected to lose $10.15

For 833 times, the expected value will be calculate as:

```{r}

n <- 833

exp_value_t <- n*exp_value

paste0("If we playes this game 833 times are expected to lose $ ", abs(exp_value_t)) # abs since we know it will be a lost

```



