---
title: "Fundamentals of Computational Mathematics - Homework 7"
author: "Jered Ataky"
date: "3/13/2021"
output:
  pdf_document: default
  html_document:
    number_sections: yes
  openintro::lab_report: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi’s. Find the distribution of Y .

## Solution

$$

 \begin{array}{l}
Y\ is\ the\ minimum\ of\ X_{i}\\
\ Suppose\ that\ each\ X_{i} \ has\ k\ possibilities:\ 1,2,...,k\ \\
\Longrightarrow \ total\ possible\ number\ of\ assignments\ for\ the\ entire\ collection\ of\ random\ \\
variables\ X_{1} ,\ X_{2} ,...,X_{n} \ is\ k^{n} .\ Thus,\ k^{n} \ is\ the\ total\ number\ of\ options.\\
Therefore,\ ( k-1)^{n} \ is\ the\ number\ of\ options\ where\ none\ of\ X_{i} \ are\ equal\ to\ 1.\\
\\
Let\ calculate\ the\ number\ of\ ways\ of\ getting\ Y:\\
\\
To\ get\ Y\ =\ 1,\ we\ will\ have\ k^{n} -( k-1)^{n} \ ways\\
To\ get\ Y\ =\ 2,\ we\ will\ get\ k^{n} -( k-2)^{n} -\left[ k^{n} -( k-1)^{n}\right] \ =\ ( k-1)^{n} -( k-2)^{n}\\
....\\
\\
In\ the\ same\ way,\ we\ can\ generalize\ it.\\
So,\ to\ get\ Y=x,\ we\ will\ have\ ( k-x+1)^{n} -( k-x)^{n} \ ways\ to\ assign\ X_{1} ,\ X_{2} ,...,X_{n} \ \\
so\ that\ the\ minimum\ value\ is\ j.\ \\
\\
We\ will\ need\ the\ to\ divide\ this\ number\ of\ ways\ by\ the\ total\ number\ of\ options\ to\ find\\
the\ distribution\ of\ Y.\ \\
\\
Therefore,\ P( Y=x) \ =\ \frac{( k-x+1)^{n} -( k-x)^{n}}{k^{n}}
\end{array}


$$


## Problem 2

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).
This machine has a manufacturer’s expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).


a. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

## Solution

```{r}

exp <- 10 # expected value

prob_exp <- 1 / exp   # Manufacturer's probability

n <- 8 # number of years

# Prob after 8 years

prob <- round(1 - pgeom(n, prob_exp), 4)

paste0("The probability that the machine will fail after 8 years is ", prob)

exp_prob <- round(1 / prob_exp, 2)

paste0("The expected value is ", exp_prob)

# Expected value before first success


std_prob <- round(sqrt(1-prob_exp) / prob_exp, 4)

paste0("The standard deviation is ", std_prob)

```



b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential. 


## Solution

```{r}

u_exp <- 10 # expected value

rate <- 1/ u_exp # rate


# Prob the machine will fail after 8 years P(x>8)

x <- 8

prob <- round(exp(-rate*x), 4) 

paste0("The probability that my MRI will fail after 8 years is ", prob)

# a. Expected value

exp_v <- 1 / rate  

paste0("Expected failure time is ", exp_v)


# Standard deviation

exp_std <- u_exp

paste0("The std is ", exp_std)



```



c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years) 


## Solution

```{r}

exp <- 10 # expected value
n <- 8 
x <- 0 # 0 success in 8 years

prob_exp <- 1 / exp   # Manufacturer's probability

prob <- round(pbinom(x, n, prob_exp), 4)

paste0("The probability that my MRI will fail after 8 years is ", prob)

# Expected value

u <- n*prob_exp

paste0("The expected value is ", u)

# Standard deviation

s <- round(sqrt(n*prob_exp*(1-prob_exp)), 4)

paste0("The standard deviation is ", s)


```

d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 


## Solution


```{r}

t <- 10
n <- 8 

x <- 0

lambda<- n / t

# After 8 years

prob <- round(ppois(x, lambda), 4)

paste0("The probability that the machine will fail after 8 years is ", prob)

# Expected value and std

u <- lambda

paste0("The expected value is ", u)

s <- round(sqrt(lambda), 4)

paste0("The standard deviation is ", s)



```



