---
title: "Fundamentals of Computational Mathematics - Homework 2"
author: "Jered Ataky"
date: "2/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem set 1


$$

 \begin{array}{l}
1.\ Proof\ and\ demonstration:\ A^{T} A\ \neq \ AA^{T}\\
\\
\\
Let\ A\ be\ a\ mXn\ matrix:\\
\\
A\ =\ \ \begin{bmatrix}
a_{11} & \cdots  & a_{1n}\\
\vdots  & \vdots \vdots \vdots  & \vdots \\
a_{m1} & \cdots  & a_{mn}
\end{bmatrix} \ \ \ \\
\\
The\ transpose\ of\ A\ is\ then\ A^{T} \ and\ given\ by:\\
\\
A^{T} \ =\ \ \ \begin{bmatrix}
a_{11} & \cdots  & a_{m1}\\
\vdots  & \vdots \vdots \vdots  & \vdots \\
a_{1n} & \cdots  & a_{mn}
\end{bmatrix}\\
\\
Let\ left\ multiply\ A\ by\ A^{T} :\\
\\
A^{T} A\ =\ \begin{bmatrix}
a_{11} & \cdots  & a_{m1}\\
\vdots  & \vdots \vdots \vdots  & \vdots \\
a_{1n} & \cdots  & a_{mn}
\end{bmatrix}\begin{bmatrix}
a_{11} & \cdots  & a_{1n}\\
\vdots  & \vdots \vdots \vdots  & \vdots \\
a_{m1} & \cdots  & a_{mn}
\end{bmatrix} \ \\
\\
=\ \begin{bmatrix}
( a_{11})( a_{11}) \ +\cdots +\ ( a_{m1})( a_{m1}) \  & \cdots  & ( a_{11})( a_{1n}) \ +\cdots +\ ( a_{m1})( a_{mn}) \ \\
\vdots  & \vdots \vdots \vdots  & \vdots \\
( a_{1n})( a_{11}) \ +\ \cdots \ +\ ( a_{mn})( a_{m1}) & \cdots  & ( a_{1n})( a_{1n}) \ +\cdots +\ ( a_{mn})( a_{mn}) \ 
\end{bmatrix} \ ( 1)\\
\\
\\
Let\ right\ multiply\ A\ by\ A^{T} :\\
\\
AA^{T} \ =\ \begin{bmatrix}
a_{11} & \cdots  & a_{1n}\\
\vdots  & \vdots \vdots \vdots  & \vdots \\
a_{m1} & \cdots  & a_{mn}
\end{bmatrix}\begin{bmatrix}
a_{11} & \cdots  & a_{m1}\\
\vdots  & \vdots \vdots \vdots  & \vdots \\
a_{1n} & \cdots  & a_{mn}
\end{bmatrix}\\
\\
=\ \begin{bmatrix}
( a_{11})( a_{11}) \ +\cdots +\ ( a_{1n})( a_{1n}) \  & \cdots  & ( a_{11})( a_{m1}) \ +\cdots +\ ( a_{1n})( a_{mn}) \ \\
\vdots  & \vdots \vdots \vdots  & \vdots \\
( a_{m1})( a_{11}) \ +\ \cdots \ +\ ( a_{mn})( a_{1n}) & \cdots  & ( a_{m1})( a_{m1}) \ +\cdots +\ ( a_{mn})( a_{mn}) \ 
\end{bmatrix} \ ( 2)\\
\\
Looking\ at\ both\ multiplication,\ \\
the\ entries\ of\ the\ matrix\ ( 1) \ are\ not\ equal\ to\ the\ entries\ of\ \\
the\ matrix\ ( 2)\\
\\
Therefore,A^{T} A\neq \ AA^{T} \ ( Proven!)\\
\\
For\ example\ the\ element\ ( 1)[ 1,1] \ =\ ( a_{11})( a_{11}) \ +\cdots +\ ( a_{m1})( a_{m1})\\
is\ different\ from\ the\ element\ ( 2)[ 1,1] \ =( a_{11})( a_{11}) \ +\cdots +\ ( a_{1n})( a_{1n}) \ \\
\\
\\
\\
\\
\end{array}

$$
Now, let consider A,  a 3X3 matrix and B the transpose of A to demonstrate the proof:


```{r}

# Given A

A <- matrix(c(2,3,1, -1,3,1, 2,-3,4), nrow = 3)
A
# Find B, transpose(A)
B <- t(A)
B
# Left multiplication:

L <- B%*%A

# Right multiplication

R <- A%*%B

```

Check if $A^{T} A\ = \ AA^{T}$

```{r}
# Check if t(A)*A = A*t(A), note that L = t(A)*A and R = A*t(A)

L == R

```

$$

As\ we\ can\ see,\ all\ the\ entries\ are\ different.
Therefore, 

A^{T} A\ \neq \ AA^{T}

$$

2. 

$$

A^{T} A\ = \ AA^{T} if\ only\ if\ A = A^{T}.

\\This\ is\ true\ if\ A\ is\ a\ symmetric\ matrix.\\

Therefore,\ we\ can\ conclude\ that\ A^{T} A\ = \ AA^{T} if\ A\ is\ a\ symmetric\ matrix.\

$$


Let consider S, a 3X3 symmetric matrix to demonstrate this:

```{r}

S <- matrix(c(1,-2,5, -2,6,8, 5,8,6), nrow = 3)

```

Compare S and its transpose

```{r}
S == t(S)

```

S == t(S). So, S is symmetric

Now, let check if  $S^{T} S\ = \ SS^{T}$

```{r}

t(S)%*%S == S%*%t(S)

```

As shown above, all the entries are the same. Therefore,  $S^{T} S\ = \ SS^{T}$
This to confirm that for a symmetric matrix S, $S^{T} S\ = \ SS^{T}$


## Problem set 2

Let consider a square matrix A less than 5X5

```{r}

lu_function <- function(A) {
  
# Let n be the dimension of the matrix 
  
  n <- dim(A)[1]
  
# Let set initially U equal to A and L the identity matrix
  
  U <- A
  L <- diag(n)
  
# Apply loop with row(i) and column(j) navigation, and find multiplier for L and U
  
  for (i in 2:n) {
    for (j in 1:(i-1)) {
      
      L[i,j] <- U[i,j] / U[j,j]
      U[i,] <- U[i,] - L[i,j]*U[j,]
    }
    
    
    
  }
  
# Return L,U
  
  return(list(L=L, U=U))
  
  
}

```


Illustration: 
Let C be a 3X3 square matrix, let factorize it and check how it works

```{r}

C <- matrix(c(1,-2,3, 4,8,4, -3,5,7), nrow = 3)

lu_function(C)


```


Compare the answer with the build-in function from matrixcalc library:


```{r echo = TRUE}

library(matrixcalc)

lu.decomposition(C)

```


Bingo!

